Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface

### 2022-03-31 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Fix flow calcs

The open split interface supplies a helium - CO~2~ mixture to a split where a capillary delivers gas to the source and the rest is vented. A flow of helium is used to displace sample CO~2~ to the delivery capillary in the split. Our open-split interface is constructed of simple, off-the-shelf parts. The core of the design is a dual needle and split housed in a VICI cross. Displacement helium flows in through one side of the cross, into the sample vial through 1/16 in. stainless steel tubing, which forms the outer flow path of the concentric dual needle. The mixture of sample CO~2~ and displacement helium flows up a needle inserted through the cross and stainless tubing which forms the inner flow path of the dual needle. The split is made from a VICI Luer adapter fitted to the outflow of the dual needle with the source delivery capillary inserted (Figure 1). The split has fairly low volume (\< 1 mL). The system was operated manually during development, but may be easily mounted on the arm of a gasbench-style autosampler to enable automated operation.

@@

The helium displacement flow rate was set to 250 μl min^-1^ for initial tests, and this was reduced to 100 μl min^-1^ for later work including most reference standard comparisons and the core comparison. The flow rate from the open split into the source is defined by 250 cm of 50 μm glass capillary (confirm this, SGE, Trajan, Agilent?), split into two sections by a zero-volume bulkhead connector at the source cage. The section inside the cage is jacketed in Teflon tubing to help protect against spark damage. The total flow rate of gas to the source varies due to the viscosity difference between CO~2~ and helium, and is estimated to be 30 μl min^-1^. ???

@@

Carbonate samples are weighed into 7 ml septa-sealed Vacutainer vials (Labco \# 768W). Sealed vials are evacuated, then 1 ml 85% phosphoric acid is added. The samples are allowed to react at 60C for 4-6 hours. Sample pressure is adjusted to 1 ATM by allowing the sample to equilibrate with a helium-filled syringe connected through the vial septum. Evacuation, acid addition, and heated reaction are performed by the automated gas handling system developed for the CFAMS GIS carbonate system [@Roberts2013].

@@

Initial work with the NEC source established optimal source parameters for consistent source performance. We used tank CO~2~ connected through a regulator to the ion source to measure source performance over a range of CO~2~ flows. Best currents for the source are around 10 μA ^12^C^-^, or roughly 1/10 of graphite target current. Best currents were produced at a CO~2~ flow rate of 7.5 μl min^-1^ CO~2~, but the source produces near optimal currents with flows from 7.5-25 μl min^-1^ CO~2~. Adding helium to the gas mixture produced similar currents to pure CO~2~ at the same CO~2~ flow rates. This is generally comparable with other hybrid sources [@Xu2007], although improved target design and continued optimization of source parameters may improve performance [@Salazar2013; @Wacker2013]. Typical currents for this study were between 4-9 μA and average current for all samples was `r curmean`. Several studies have shown best currents at lower CO~2~ flow rates [@Xu2007; @Ruff2010]. While efficiency of the source is best at very low flow rates (\<1.5 μl min^-1^), measuring samples with flow set for optimal current produces more precise results due to higher count rates.

>>>>

The open split interface supplies a helium - CO~2~ mixture to a split where a capillary delivers gas to the source and the rest is vented. A flow of helium is used to displace sample CO~2~ to the delivery capillary in the split. Our open-split interface is constructed of simple, off-the-shelf parts. The core of the design is a dual needle and split housed in a VICI cross. Displacement helium flows in through one side of the cross, into the sample vial through 1/16 in. stainless steel tubing, which forms the outer flow path of the concentric dual needle. The mixture of sample CO~2~ and displacement helium flows up a needle inserted through the cross and stainless tubing which forms the inner flow path of the dual needle. The split is made from a VICI Luer adapter fitted to the outflow of the dual needle with the source delivery capillary inserted (Figure 1). The split has fairly low volume (< 1 mL). The system was operated manually during development, but may be easily mounted on the arm of a gasbench-style autosampler to enable automated operation.

@@

```{r}
# Flow calc for CO2
co2flow <- flowcalc(100, 2.5E-5, 1.47E-5, 1.50)

# Flow calc for helium
heflow <- flowcalc(101, 2.5E-5, 1.96E-5, 1.50)
```

The helium displacement flow rate was set to 250 μl min^-1^ for initial tests, and this was reduced to 100 μl min^-1^ for later work, including most reference standard comparisons and the core comparison. The flow rate from the open split into the source is defined by 1.5 m of 50 μm glass capillary (confirm this, SGE, Trajan, Agilent?), split into two sections by a zero-volume bulkhead connector at the source cage. The section inside the cage is jacketed in Teflon tubing to help protect against spark damage. The total flow rate of gas to the source varies due to the viscosity difference between helium and CO~2~ and helium, and is estimated to be between 30 μl min^-1^ for pure helium and 40 μl min^-1^ for pure CO~2~ using the Hagen-Poiseuille equation.

@@

Carbonate samples are weighed into 7 ml septa-sealed Vacutainer vials (Labco 768W). Sealed vials are evacuated, then 1 ml 85% phosphoric acid is added. The samples are allowed to react at 60C for 4-6 hours. Sample pressure is adjusted to 1 ATM by allowing the sample to equilibrate with a helium-filled syringe connected through the vial septum. Evacuation, acid addition, and heated reaction are performed by the automated gas handling system developed for the CFAMS GIS carbonate system [@Roberts2013].

@@

Initial work with the NEC source established optimal source parameters for consistent source performance. We used tank CO~2~ connected through a regulator to the ion source to measure source performance over a range of CO~2~ flows. Best currents for the source are around 10 μA ^12^C^-^, or roughly 1/10 of graphite target current. Best currents were produced at a CO~2~ flow rate of 7.5 μl min^-1^ CO~2~, but the source produces near optimal currents with flows from 7.5-25 μl min^-1^ CO~2~. Adding helium to the gas mixture produced similar currents to pure CO~2~ at the same CO~2~ flow rates. This is generally comparable with other hybrid sources [@Xu2007], although improved target design and continued optimization of source parameters may improve performance [@Salazar2013; @wacker_versatile_2013]. Typical currents for this study were between 4-9 μA and average current for all samples was `r curmean`. Several studies have shown best currents at lower CO~2~ flow rates [@Xu2007; @Ruff2010]. While efficiency of the source is best at very low flow rates (\<1.5 μl min^-1^), measuring samples with flow set for optimal current produces more precise results due to higher count rates.
